Measuring individual differences in the understanding of gaze cues

### 2022-05-20 paper/manuscript.Rmd
first draft intro

Maybe for abstract: Developmental psychology is facing a dilemma: many research questions are questions about individual differences, yet, there is a lack of tasks to reliably measure these individual differences.

# Introduction 
Social cognition – representing and reasoning about an agent’s perspectives, knowledge states, intentions, beliefs, and preferences to explain and predict their behavior – is among the best-studied phenomena in developmental research.

In recent decades, much progress has been made in determining the average age at which a specific social-cognitive ability emerges in development.

But we still know relatively little about the co-development of social-cognitive abilities, variation in the individual’s pace of mastering developmental milestones, and which factors influence a child’s proficiency.

Researchers are interested in questions concerning individual differences in the social cognition domain [@hughes2015individual]. Previous studies focused, for example, on the association of stereotypes and theory of mind [@rizzo2018theory], mental state talk and false belief understanding [@hughes2011individual], attachment quality and gaze following [@astor2020social], language and gaze following [@okumura2017individual], emotion, language, family interaction and belief understanding [@dunn1991young; @cutting1999theory; @bulgarelli2016social], and the interplay of physical and social cognition [@herrmann2010structure]. Correlational approaches have also been used to argue for and test assumptions of cognitive development theories [@mundy2007individual; @kidd2018individual; @underwood1975individual].

This is due to the fact that many developmental psychology studies track change at an (age-) group level. 
Traditional measures of social cognition are not designed to capture variation between children: 
they often rely on low trial numbers, small sample sizes, dichotomous measures, and consequently lack satisfactory psychometric properties. 
This has profound implications for what these studies can show. Poor measurement of social cognition on an individual level may conceal relations between different aspects of cognition and may obscure developmental change.

To fully understand how social-cognitive abilities emerge and relate to each other, we must take individual differences seriously.

We need tasks that capture variability in performance reliably (i.e., systematically order individuals in the same way) and meaningfully (i.e., differences in test scores correspond to differences in ability).

A recent paper by @hedge2018reliability argued that cognitive tasks often become well-established by displaying robust effects across studies. The authors could show that this often does not emerge due to high measurement variance but rather low between-subject variability, concluding that popular cognitive tasks may reliably measure group differences but not individual differences. In a similar vein, @pronk2021methods reasoned that accurate reliability estimates are needed to judge how well a certain cognitive test is suited to draw inferences about individuals.

[TODO: check Manuels references]

However, we rarely find studies addressing these issues. 
Methods, that have currently been used, are...
One of the most commonly applied measures in social cognition research is the Sally-Anne task. [TODO: fill in FB studies]
The outcome measures false belief understanding in a dichotomous way: children pass the task if they answer that the protagonist would ...
With four to five years of age, most children pass the task. From this age onwards, the change-of-location task shows ceiling effects and has very limited diagnostic value. With these procedures, researchers could systematically determine the average age at which children succeed in the Sally-Anne task. However, no individual differences can be visible, especially when applying only one trial per child.

Our aim is, therefore, to make individual differences measurable in a systematic and reliable way. For getting precise person-specific estimates, we need an accurate measure of a child's given underlying social-cognitive ability. One of our main goals is to create a task that induces variation between children. To approach this issue, we focus on a continuous measure and short trials that facilitate more than a dozen replicates per subject.

To estimate the effect of inter-individual differences, we need a large sample size since some of the effects that we are looking for might be small. Since data collection with families requires a lot of organizational effort, we want to enable data collection in-person with supervision as well as remotely without supervision. A standardized, easily accessible test procedure helps us to enable data collection at scale.

Another goal in creating new tasks should be to focus on the face value: we want to measure the underlying social-cognitive ability as straight-forward and directly as possible. Importantly, we can assess whether the task shows external validity by assessing the relationship with related constructs and factors of children's social surroundings.

In our first novel task, we concentrated on the fundamental ability that is involved in many social-cognitive reasoning processes: gaze cue understanding - the ability to locate and use the attentional focus of an agent. The first component is often termed gaze following - turning ones eyes in the same direction as the gaze of another agent - and has been studied intensively [TODO: add references]. 
By acting as a 'front end ability' [@brooks2005development, p. 535], following an agent's gaze provides insights into their intentions, thoughts, and feelings. 
In our definition, gaze cue understanding goes one step further by including the *acting* on the gaze-cued location.

---

- why is social cognition important
- what methods are currently been used?: wellman
- what are common issues? 
- what to aim at. individual differences in developmental psychology
- what characteristics should a new task fulfill? reliable tasks, variation needed, more trials
- goal of the current project: standardized, easy to use continuous methods
- soc cog, gaze cue understanding small def (longer one dev paper)

---

Rakoczy, H. (2022). Foundations of theory of mind and its development in early childhood. Nature Reviews Psychology, 1–13. https://doi.org/10.1038/s44159-022-00037-z:

"ToM also has specific real-life consequences. First, the development of ToM competence goes along with general measures of children’s peer social skills in early and middle childhood. (...) Second, ToM specifically predicts communicative competence. (...) Third, ToM competence is related to the quality of peer relationships: children with more advanced ToM are rated as more likeable and popular among their peers. Fourth, children who are more proficient at ToM tasks tend to act more prosocially, including comforting, sharing or helping other individuals. Finally, preschool ToM competence predicts achievement in primary school, a relationship that is possibly mediated by social competence, in that preschool ToM abilities enable subsequent social competence development, which in turn contributes to school achievement." (p. 2)
"Evidence for an emerging understanding of perception at 9 months of age comes from various sources. For example, children begin to follow the gaze of other agents in systematic and differential ways: they follow an agent’s head turn only when the agent can actually see (has their eyes open rather than closed, or wears a transparent rather than an opaque blindfold)." (p. 2)
Developmental determinants: executive function, language ("that" complementations), social (SES, siblings, mind-minded parents)
Implicit tasks: "A third class of implicit ToM tasks is interaction tasks, in which participants are involved in a communicative or cooperative interaction with another agent. This agent forms a mental state (such as a true or false belief regarding the contents of a box) and experimenters measure whether participants spontaneously take the agent’s belief into account in their interaction with the agent (for instance by helping or by interpreting the agent’s communicative acts accordingly)" (p. 9)
=> reliable & valid tasks to assess coherent development of social-cognitive functions

>>>>

# Introduction 
<!-- ## What is Social Cognition, and why do we care? -->
Social cognition – representing and reasoning about an agent’s perspectives, knowledge states, intentions, beliefs, and preferences to explain and predict their behavior – is among the best-studied phenomena in developmental research. 
In recent decades, much progress has been made in determining the average age at which a specific social-cognitive ability emerges in development [@gopnik1991young; @wellman2001metaanalysisa; @wellman2004scalinga; @peterson2012mind; @rakoczy2022foundations].

<!-- ## How has it been measured traditionally? -->
One of the most commonly applied, prototypical measures for social cognition is the Sally-Anne task [@baron-cohen1985does]. Here, children watch a short sequence of events (often acted out or narrated by the experimenters). A doll called Sally puts her marble into a basket. After Sally leaves the scene, a second doll named Anne takes the marble and moves it into a box. Participants then get asked where Sally will look for her marble once she returns. The outcome measures false belief understanding in a dichotomous way: children pass the task if they take the protagonist’s epistemic state into account and answer that she will look into the basket. Many years of research utilizing these verbal change-of-location tasks suggest that children develop belief-representing abilities at four to five years of age [for a review, see @wellman2001metaanalysisa]. Several cross-cultural studies supported this evidence [@barrett2013early; @callaghan2005synchrony; cf. @mayer2015weird].
However, from this age onwards, the change-of-location task shows ceiling effects and has very limited diagnostic value. As @wellman2012theory put it, “it’s really only passing/failing one sort of understanding averaged across age” (p. 317). Therefore, the focus has been to track change at an (age-) group level – within a narrow developmental time frame and with a strong emphasis on a single social-cognitive ability.

<!-- ## Why care about individual differences? -->
Although humans substantially differ in their behavior, thoughts, and motivations, researchers often regard variability as cumbersome or reinterpret it as error variance, therefore neglecting and trivializing individual differences [@kidd2018individual]. Consequently, little consensus has been reached on the interrelation or independence of social-cognitive abilities [termed ‘structure of social cognition’ by @happe2017structure].

However, identifying variability in social-cognitive abilities and factors influencing those could be vital in theory building (e.g., to test causal predictions) and design interventions [@happe2017structure]. For example, it could be used to provide evidence for accounts on *developmental lags* (i.e., whether children differ in their speed of acquiring new conceptual insights into the understanding of the mind) or *genuine variation* (i.e., whether children differ in their ease or fluency in (applying) these social-cognitive abilities) [@hughes2015individual]. 
Therefore, we argue that we should study the co-development of social-cognitive abilities, variation in the individual’s pace of mastering developmental milestones, and which factors influence a child’s proficiency. For this endeavor, we need to take individual differences seriously – both theoretically and methodologically.

<!-- ## Existing research tapping into individual differences -->
Some studies have already examined individual differences in social cognition [@hughes2015individual; @slaughter2015theorya]. Previous studies focused, for example, on the association of mental state talk and false belief understanding [@hughes2011individual], emotion, language, family interaction and belief understanding [@dunn1991young; @cutting1999theory; @bulgarelli2016social], stereotypes and theory of mind [@rizzo2018theory], prosociality and theory of mind [for a review see @imuta2016theory; @walker2005gender], executive functions and theory of mind [@carlson2004individual; @buttelmann2021relations], parenting practices and theory of mind [for a review see @pavarini2013parental], agency and moral cognition [@sodian2016understanding], attachment quality and gaze following [@astor2020social], language and gaze following [@okumura2017individual], and the interplay of physical and social cognition [@herrmann2010structure]. 
Correlational approaches have also been used to argue for and test assumptions of cognitive development theories [@mundy2007individual; @kidd2018individual; @underwood1975individual].

<!-- ## Methodological limitations -->
Frequently, developmental psychologists are surprised to find minor or no association between constructs that are thought to be theoretically related [e.g., @sodian2016understanding]. This might be because traditional measures of social cognition are not designed to capture variation between children: they often rely on low trial numbers, small sample sizes, and dichotomous measures. 
@hedge2018reliability argued that cognitive tasks often become well-established by displaying robust effects across studies. The authors could show that this usually does not emerge due to high measurement variance but relatively low between-subject variability, concluding that popular cognitive tasks may reliably measure group differences but not individual differences. Similarly, @pronk2021methods reasoned that accurate reliability estimates are needed to judge how well a particular cognitive test is suited to draw inferences about individuals.
A recent review underlined reoccurring issues in standardization, validity, reliability, and sensitivity to developmental change: alarmingly few theory of mind measures reported empirically validated psychometric properties [@beaudoin2020systematic] or show poor test-retest reliability [@mayes1996testretest].

Even studies focusing on individual differences often use behavioral forced-choice paradigms [e.g., carlson2004individual; @walker2005gender]. Few studies that do realize continuous measures often administer (theory of mind) scales and conventionally compute aggregate scores (across tasks or trials), therefore neglecting the issue of measurement error [@hughes2011individual]. This way, a particular social-cognitive ability in question is still assessed categorically on a trial basis [e.g., @bulgarelli2016social; @rizzo2018theory; @buttelmann2021relations]. 
This has profound implications for what these studies can show. Poor measurement of social cognition on an individual level may conceal relations between different aspects of cognition and may obscure developmental change.

<!-- ## A wish list for new social-cognitive tasks -->
Therefore, developmental psychology is facing a dilemma: many research questions are questions about individual differences, yet, there is a lack of tasks to reliably measure these individual differences. To capture the emergence of social-cognitive abilities and related social factors in greater precision and detail, we must consequently address the methodological limitations of existing study designs [@hughes2004links; @hughes2011individual].

@schaafsma2015deconstructing compiled a “wish-list” for new social-cognitive paradigms and advocated for parametric measures covering proficiency as a range, avoiding floor and ceiling effects, and showing satisfactory test-retest reliability estimates. In valid and reliable tasks, differences in test scores should correspond to differences in the social-cognitive ability and systematically order individuals in the same way. As a nice add-on, tasks should allow for use within an fMRI scanner. With an easily administrable test procedure, larger sample sizes can be collected to estimate the potentially small effect of inter-individual differences.

Additionally, it has been argued that social-cognitive abilities should not be treated as an all-or-nothing matter (e.g., dichotomously in pass/fail situations) but as abilities on a continuum [@hughes2015individual; @beaudoin2020systematic]. Task batteries might be better suited to cover a range of social-cognitive abilities [@hughes2011individual]. A proposal was that first basic social-cognitive processes should be relatively well understood before advancing to more complex social-cognitive abilities [@schaafsma2015deconstructing].

Another goal in creating new tasks should be to focus on the face value: we want to measure the underlying social-cognitive ability as straight-forward and directly as possible. The task should serve as a proxy for behavior as it appears in the real world. Importantly, we can assess whether the task shows external validity by evaluating the relationship with related constructs and factors of children's social surroundings.

<!-- ## Our aims -->
Our aim is, therefore, to make individual differences measurable in a systematic and reliable way. We want to create a task that induces variation between children to get precise person-specific estimates. To approach this issue, we focus on a continuous measure and short trials that facilitate more than a dozen replicates per subject. A standardized, easily accessible test procedure helps us to enable data collection at scale. Since data collection with families requires a lot of organizational effort, we want to enable data collection in-person with supervision as well as remotely without supervision.

In our first novel task, we concentrate on the fundamental ability that is involved in many social-cognitive reasoning processes: gaze cue understanding - the ability to locate and use the attentional focus of an agent. The first component of this ability is often termed gaze following - turning one's eyes in the same direction as the gaze of another agent - and has been studied intensively [@itakura1998use; @delbianco2019developmental; @byers-heinlein2021development; @moore2008development; @coelho2006searching; @tomasello2007reliance; @hernik2019infant; @astor2021gaze; @frischen2007gaze; @shepherd2010following; @lee1998children]. Following an agent's gaze provides insights into their intentions, thoughts, and feelings by acting as a 'front end ability' [@brooks2005development, p. 535]. 
In our definition, gaze cue understanding goes one step further by including the *acting on the gaze-cued location* - therefore, using the available social information to guide your behavior as needed in real-life conditions.
